---
title: "Homework 03"
Author: "Carla Quintanilla"
Date: "05-22-25"
format: html
---

[Github Repositopry] https://github.com/carlaaa02/ENVS-193DS_homework-03

## Set up

```{r}
#| message: false

#Loading Packages 
library(tidyverse)
library(here)
library(gt)
library(flextable)
library(janitor)
library(readxl)

#Read in data 
data <- read_csv(here("data","my_data.csv"))

```

## Problem 1 

a. I could calculate the mean focus duration to compare how long I stayed focused between different background noise levels (e.g., quiet, semi-loud, loud), because I suspect that noise distractions in the library or living room affect my ability to concentrate. This comparison would help me understand whether quieter environments are actually leading to longer periods of focus.

b. Visualization

```{r}
clean_data <- data |> 
  clean_names() |> 
   mutate(
    background_noise = str_trim(background_noise),
    background_noise = factor(background_noise, levels = c("quite", "semi-loud", "loud"))
  )

ggplot(data, aes(x = background_noise, y = focus_duration_hrs, fill = background_noise)) +
  geom_jitter(width = 0.2, color = "black", size = 2, alpha = 0.7) +
  stat_summary(fun = mean, geom = "bar", width = 0.5, alpha = 0.6, color = "black") +
  scale_fill_manual(values = c("quite" = "#1f78b4", "semi-loud" = "#33a02c", "loud" = "#e31a1c")) +
  labs(
    title = "Average Focus Duration by Background Noise",
    x = "Background Noise Level",
    y = "Focus Duration (hours)",
    fill = "Noise Level"
  ) +
  theme_minimal(base_size = 14)
```

c. Caption 

This figure shows the relationship between background noise level and focus duration (in hours). Individual observations are jittered, while bars represent the average focus time for each noise category. The data suggest that quieter environments may support longer periods of focused work.

d. Table Representation 

```{r}

summary_table <- clean_data |> 
  group_by(background_noise) |> 
  summarise(mean_focus = round(mean(focus_duration_hrs),1)) |> 
  gt() |>
  tab_header(
    title = "Mean Focus Duration by Background Noise Level"
  ) |>
  cols_label(
    background_noise = "Background Noise Level",
    mean_focus = "Mean Focus Duration (hrs)"
  )

summary_table 
```




